---
title: "PRIVILEGED AND CONFIDENTIAL: Ethanol Growth Report"
author: "Geoffrey Zahn"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE)


library(tidyverse)
library(easystats)
library(patchwork)
library(broom)
'%ni%' <- Negate('%in%')
theme_set(theme_minimal())

df <- read_csv("./data/colony_counts.csv")
```


# **Protocol:**

**Replication of Ewaze, *et al.*, 2008, Figure 4 - "Effect of ethanol vapour on dormant cell germination"**

Ewaze, J. O., Summerbell, R. C., & Scott, J. A. (2008). Ethanol physiology in the warehouse-staining fungus, Baudoinia compniacensis. Mycological Research, 112(11), 1373–1380. https://doi.org/10.1016/j.mycres.2008.05.003

**Design**

6 Treatments (Sterile control, Acetone 10ppm control, Ethanol 0.1ppm, Ethanol 1ppm, Ethanol 10ppm, Ethanol 100ppm)

5 replicate subjects per treatment

Grow dormant colonies on sterile filters for 14 days in vapor treatment chambers

Count resultant colonies on filters

Analyze with One-way ANOVA and GLM (Colony count ~ treatment)

___

```{r}
# find control group measures
H2O_control <- df %>% 
  filter(ppm == 0) %>% 
  summarize(H2O_mean=mean(colony_count),
            H2O_sd=sd(colony_count),
            H2O_upper=H2O_mean + H2O_sd,
            H2O_lower=H2O_mean - H2O_sd)
# Acetone_control <- df %>% 
#   filter(Treatment == "Acetone") %>% 
#   summarize(Acetone_mean=mean(CFUs),
#             Acetone_sd=sd(CFUs),
#             Acetone_upper=Acetone_mean + Acetone_sd,
#             Acetone_lower=Acetone_mean - Acetone_sd)
controls <- H2O_control
```


```{r}
# build plots
etoh <- df %>% 
  filter(ppm != 0) %>% 
  # filter(colony_count < 200) %>% 
  mutate(Concentration_ppm = factor(ppm,levels = c("0.1","1","10","100")))

h2o_plot <- 
etoh %>%
  ggplot(aes(x=Concentration_ppm,y=100*(colony_count/H2O_control$H2O_mean),fill=Concentration_ppm)) +
  geom_boxplot() +
  labs(x="Ethanol conc. (ppm)",
       title="Post-exposure germination\nvs water control (%)",
       y=NULL) +
  geom_hline(yintercept=100*H2O_control$H2O_mean/H2O_control$H2O_mean) +
  geom_hline(yintercept=100*H2O_control$H2O_upper/H2O_control$H2O_mean, linetype=2) +
  geom_hline(yintercept=100*H2O_control$H2O_lower/H2O_control$H2O_mean, linetype=2) +
  coord_cartesian(ylim=c(0,1000)) +
  # annotate("text",x=2.5,y=107,label="Water control values") +
  theme(axis.text = element_text(face='bold',size=12),
        legend.position = 'none')

```


```{r}
# run stats
etoh <- etoh %>% 
  mutate(relative_growth_h2o = 100 * colony_count / controls$H2O_mean)


water_cfus <- df %>% 
  filter(ppm == 0) %>% 
  pluck("colony_count")
cfus_0.1 <- df %>% 
  filter(ppm == "0.1") %>% 
  pluck("colony_count")
cfus_1 <- df %>% 
  filter(ppm == "1") %>% 
  pluck("colony_count")
cfus_10 <- df %>% 
  filter(ppm == "10") %>% 
  pluck("colony_count")
cfus_100 <- df %>% 
  filter(ppm == "100") %>% 
  pluck("colony_count")

# water t.tests
water_ttests <- list()
for(i in ls(pattern = "cfus_")){
  d <- get(i)
  water_ttests[[i]] <- t.test(d,water_cfus) %>% tidy()
}



```

**Short summary:**

  - While it appears that ethanol at 1, 10, and 100 ppm increases *B. panamericana* germination, statistical tests are unable to conclude that these differences are significant and real.
  - With the exception of a single 100 ppm jar which had 262 colonies growing, the remainder of samples had between 0 and 20 colonies.
  - I have left that high-count jar in all analyses, but I am suspicious of why it did so much better than every other 100 ppm sample. It could be contamination with another carbon source, or just that biology is confusing.
  
___

# **Results:**


```{r}

# add meaningful fill color to plots
fillcolors <- data.frame(
Water_Diff = water_ttests %>% 
  map(5) %>% 
    map_lgl(function(x){x<0.05})) %>% 
  mutate(PPM=row.names(.)) %>% 
  select(PPM,Water_Diff) %>% 
  mutate(PPM=PPM %>% str_remove("cfus_")) %>% 
  mutate(watercolor=case_when(Water_Diff == TRUE ~ "DarkGreen",
                              TRUE ~ "Gray"))

# adapt plots to have meaningful fill colors

h2o_plot <- h2o_plot + 
  scale_fill_manual(values = fillcolors$watercolor)
h2o_plot
  

  


```

Figure 1.

Post-exposure germination values shown as percentages of sterile water and acetone control germination values. Control germination ranges shown in horizontal zones around 100%. Green boxes indicate germination values that differed significantly from the respective control values.

<br>

___

I used several complimentary (redundant) statistical approaches to determine whether apparent differences (from Figure 1) were statistically significant. First, I ran a General Linear Regression on raw colony count data, using Ethanol concentration as a predictor. There was no significant difference between the treatments.


<br>

```{r tidy=FALSE, results = 'asis'}
df %>% 
  glm(formula = colony_count ~ factor(ppm),
      data=.) %>% 
  report() %>% 
  summary()
  

```

Next, I ran the alternative approach of conducting individual T-tests between each Ethanol treatment and the sterile water controls. This series of tests agreed with the GLM approach.

```{r}

data.frame(
Water_Diff = water_ttests %>% 
  map(5) %>% 
    map_lgl(function(x){x<0.05})) %>% 
  mutate(PPM=row.names(.)) %>% 
  select(PPM,Water_Diff) %>% 
  mutate(PPM=PPM %>% str_remove("cfus_")) %>% 
  remove_rownames() %>% 
  kableExtra::kable() %>% 
  kableExtra::kable_classic(lightable_options = c('hover')) 

```

Table 1 - Shows whether there was a statistically significant difference between observed growth and controls (both water and acetone) for each concentration of Ethanol. Paired T-tests were conducted between each ethanol concentration and the respective control.


<br>

Lastly, I ran another linear regression model, but this time on transformed values instead of raw colony counts. In this case, relative colony counts were used (relative to the sterile water control mean colony count value). This was to better reflect what was visualized in Figure 1. Again, the model agreed that it is unable to determine significant differences between the different Ethanol treatments in their relative increase in germination over sterile water controls.

```{r tidy=FALSE, , results = 'asis'}
df %>% 
  mutate(Relative_germination_water = colony_count / controls$H2O_mean) %>% 
  # mutate(Concentration_ppm = factor(Concentration_ppm,levels = c("0.1","1","10","100"))) %>% 
  pivot_longer(starts_with("Relative_"),
               names_to = "Control",
               names_prefix = "Relative_germination_",
               values_to = "Relative_germination") %>% 
  filter(ppm > 0) %>% 
  glm(formula = Relative_germination ~ ppm,
      data = .) %>% 
  report() %>% 
  summary()
  

```